DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-04-20 paper/conclusion.tex
fixes and acknowledgement

Despite the complexity of the neural network, DeepPicar uses a
low-cost embedded quad-core computer, the Raspberry Pi 3,  as its sole
computing resource.
We systematically analyzed the platform's real-time capability in
supporting the CNN-based real-time control task.
We also evaluated other, more powerful, embedded computing
platforms to better understand achievable real-time performance of
DeepPicar's CNN based control system and the impact of
computing hardware architectures.
We find all tested embedded platforms, including the Pi 3, are capable
of supporting deep-learning based real-time control, from 20 Hz up to
100 Hz, depending on the platform and its system
configuration. Futhermore, all platforms were capable of consolidating
multiple CNN models and/or tasks.

However, when consolidating multiple models/tasks, shared resource
contention remains an important issue that must be considered to
ensure desired real-time performance on these shared memory based
embedded computing platforms. 
Toward this end, we evaluated the impact of shared resource contention
to the CNN workload in diverse consolidated workload setups, and
evaluated the effectivness of state-of-the-art shared resource
isolation mechanisms in protecting real-time performance of the CNN
workload.

>>>>

Despite the complexity of the neural network, DeepPicar uses a
low-cost embedded quad-core computer, the Raspberry Pi 3,  as its sole
computing resource.
We systematically analyzed the platform's real-time capability in
supporting the CNN-based real-time control task.
We also evaluated other, more powerful, embedded computing
platforms to better understand achievable real-time performance of
DeepPicar's CNN based control system and the impact of
computing hardware architectures.
We find all tested embedded platforms, including the Pi 3, are capable
of supporting the CNN based real-time control, from 20 Hz up to
100 Hz, depending on the platform.
Futhermore, all platforms were capable of consolidating
multiple CNN models and/or tasks.

However, we also find that shared resource
contention remains an important issue that must be considered to
ensure desired real-time performance on these shared memory based
embedded computing platforms.
Toward this end, we evaluated the impact of shared resource contention
to the CNN workload in diverse consolidated workload setups, and
evaluated the effectivness of state-of-the-art shared resource
isolation mechanisms in protecting performance of the CNN
based real-time control workload.

### 2018-04-20 paper/evaluation-cachepart.tex
fixes and acknowledgement

Cache partitioning is a well-known technique to improve isolation in
multicore systems by giving a dedicated cache space to each individaul 
task or core. In this experiment, we use PALLOC~\cite{yun2014rtas}, a
page-coloring based kernel-level memory allocator for Linux.
Page coloring is a OS technique that controls the physical addresses
of memory pages. By allocating pages over non-overlapping cache sets,
the OS can effectively partition the cache.
Using PALLOC, we investigate the effect of cache partitioning on
protecting DeepPicar's CNN based controller.

>>>>

Cache partitioning is a well-known technique to improve isolation in
a multicore system by giving a dedicated cache space to each individaul 
task or core. In this experiment, we use PALLOC~\cite{yun2014rtas}, a
page-coloring based kernel-level memory allocator for Linux.
Page coloring is a OS technique that controls the physical addresses
of memory pages. By allocating pages over non-overlapping cache sets,
the OS can effectively partition the cache.
Using PALLOC, we investigate the effect of cache partitioning on
protecting DeepPicar's CNN based controller.

### 2018-04-20 paper/evaluation.tex
fixes and acknowledgement

In this section, we experimentally analyze various real-time aspects
of the DeepPicar. This includes
(1) measurement based worst-case execution time (WCET) analysis of
deep learning inferencing,
(2) the effect of using multiple cores in accelerating inferencing,
(3) the effect of co-scheduling multiple deep neural network models,
and 
(4) the effect of co-scheduling memory bandwidth intensive co-runners.
%% and
%% (5) the effect of partitioning the shared L2 cache.

>>>>

In this section, we experimentally analyze various real-time aspects
of the DeepPicar. This includes
(1) measurement based worst-case execution time (WCET) analysis of
deep learning inferencing,
(2) the effect of using multiple cores in accelerating inferencing,
(3) the effect of co-scheduling multiple deep neural network models,
and 
(4) the effect of co-scheduling memory bandwidth intensive co-runners,
and
(5) the effect of shared L2 cache partitioning and memory bandwidth
throttling for guaranteed real-time performance.

### 2018-04-20 paper/main.tex
revision up to Section 4.E
fixes and acknowledgement

\author{Michael G. Bechtel$^\dagger$, Elise McEllhiney$^\dagger$, Minje Kim$^\star$, Heechul Yun$^\dagger$\\
  $^\dagger$ University of Kansas, USA. \{mbechtel, e908m428, heechul.yun\}@ku.edu \\
  $^\star$ Indiana University, USA. minje@indiana.edu \\
}

@@

\begin{abstract}
We present DeepPicar, a low-cost deep neural network (DNN) based
autonomous car platform. DeepPicar is a small scale
replication of a real self-driving car called Dave-2 by NVIDIA, which
drove on public roads using a deep convolutional neural network (CNN), 
that takes images from a front-facing camera as input and produces
car steering angles as output. DeepPicar uses almost the exact same 
network architecture---9 layers, 27 million connections and 250K
parameters---and can be trained to drive itself, in real-time, using a
web camera and a modest Raspberry Pi 3 quad-core platform.
Using DeepPicar, we analyze the Pi 3's computing capabilities to 
support end-to-end deep learning based real-time control of autonomous
vehicles. We also systematically compare other contemporary embedded
computing platforms using the DeepPicar's CNN based real-time control
software as a workload. 
We find all tested platforms, including the Pi 3, are capable of
supporting deep-learning based real-time control, from 20 Hz up to 100
Hz depending on hardware platform. 
However, shared resource contention remains an
important issue that must be considered in applying deep-learning
models on shared memory based embedded computing platforms.
\end{abstract}

@@

%-------------------------------------------------------------------------
\input{intro}
\input{background}
\input{overview}
\input{evaluation}
%\input{evaluation-cachepart}
\input{comparison}
%\input{discussion}
\input{related}
\input{conclusion}
%-------------------------------------------------------------------------
\bibliographystyle{abbrv}
\bibliography{reference}

>>>>

\author{Michael G. Bechtel$^\dagger$, Elise McEllhiney$^\dagger$, Minje Kim$^\star$, Heechul Yun$^\dagger$\\
  $^\dagger$ University of Kansas, USA. \{mbechtel, elisemmc, heechul.yun\}@ku.edu \\
  $^\star$ Indiana University, USA. minje@indiana.edu \\
}

@@

\begin{abstract}
We present DeepPicar, a low-cost deep neural network based
autonomous car platform. DeepPicar is a small scale
replication of a real self-driving car called DAVE-2 by NVIDIA. DAVE-2
drove on public roads using a deep convolutional neural network (CNN), 
which takes images from a front-facing camera as input and produces
car steering angles as output. DeepPicar uses the same 
network architecture---9 layers, 27 million connections and 250K
parameters---and can be trained to drive itself using a
web camera and a modest Raspberry Pi 3 quad-core platform in real-time.
Using DeepPicar, we analyze the Pi 3's computing capabilities to 
support end-to-end deep learning based real-time control of autonomous
vehicles. 
We also systematically compare other contemporary embedded
computing platforms using the DeepPicar's CNN based real-time control
workload.

We find all tested platforms, including the Pi 3, are capable of
supporting the CNN based real-time control, from 20 Hz up to 100
Hz depending on hardware platform. 
However, shared resource contention remains an
important issue that must be considered in applying deep-learning
models on shared memory based embedded computing platforms; we observe
up to 11.6X control loop execution time increase due to such shared
resource contention. We evaluate state-of-the-art cache
partitioning and memory bandwidth throttling techniques on the Pi 3
in protecting the CNN workload. We find cache partition is
ineffective, while memory bandwidth throttling is an effective
solution.
\end{abstract}

@@

%-------------------------------------------------------------------------
\input{intro}
\input{background}
\input{overview}
\input{evaluation}
%\input{evaluation-cachepart}
\input{comparison}
%\input{discussion}
\input{related}
\input{conclusion}
\input{acknowledgement}
%-------------------------------------------------------------------------
\bibliographystyle{abbrv}
\bibliography{reference}

### 2018-04-20 paper/overview.tex
fixes and acknowledgement

\begin{figure}[h]
  \centering
  \includegraphics[width=.4\textwidth]{figs/DeepPicar_platform}
  \caption{DeepPicar platform.}
  \label{fig:overview}
\end{figure}

@@

Note that although the steering angle output of the network ($angle$) is
a continuous real value, the RC car platform we used only supports
three discrete angles---left (-30$^{\circ}$), center 
(0$^{\circ}$), and right (+30$^{\circ}$)---as control inputs.
We approximate the network generated real-valued angle to the closest
one of the three angles. Although this may introduce inaccuracy in
control, the observed control performance of the system is respectable,
likely due to the inherently stocastic nature of CNNs.

@@

Our trained CNN models showed good prediction accuracy, successfully
navigating several different tracks we used for training.
For instance, the DeepPicar could remain on Track 1
(Figure \label{fig:track}) over 10 minutes at a moderate speed (50\%
throttle), at which point we stopped the experiment, and more than one
minute at a higher speed (75\% throttle)~\footnote{Self-driving videos: \url{https://photos.app.goo.gl/q40QFieD5iI9yXU42}
  %% \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}
}. Running at
higher speed is inherently more challenging because the CNN controller
has less time to recover from mistakes (bad predictions).  Also, we
find that the prediction accuracy is significantly affected by the
quality of training data as well as various environmental factors such
as lighting conditions. We plan to investigate more systematic ways
to improve the CNN's prediction accuracies.

>>>>

\begin{figure}[h]
  \centering
  \includegraphics[width=.38\textwidth]{figs/DeepPicar_platform}
  \caption{DeepPicar platform.}
  \label{fig:overview}
\end{figure}

@@

Note that although the steering angle output of the CNN ($angle$) is
a continuous real value, the RC car we used only supports
three discrete angles---left (-30$^{\circ}$), center 
(0$^{\circ}$), and right (+30$^{\circ}$)---as control inputs.
We approximate the network generated real-valued angle to the closest
one of the three angles. Although this may introduce inaccuracy in
control, the observed control performance of the system is respectable,
likely due to the inherently stocastic nature of CNNs.

@@

Our trained CNN models showed good prediction accuracy, successfully
navigating several different tracks we used for training.
For instance, the DeepPicar could remain on Track 1
(Figure~\ref{fig:track}) over 10 minutes at a moderate speed (50\%
throttle), at which point we stopped the experiment, and more than one
minute at a higher speed (75\% throttle)~\footnote{Self-driving videos: \url{https://photos.app.goo.gl/q40QFieD5iI9yXU42}
  %% \url{https://photos.app.goo.gl/ce93sU7jPk4ywO8u2}
}. Running at
higher speed is inherently more challenging because the CNN controller
has less time to recover from mistakes (bad predictions).  Also, we
find that the prediction accuracy is significantly affected by the
quality of training data as well as various environmental factors such
as lighting conditions. We plan to investigate more systematic ways
to improve the CNN's prediction accuracies.

### 2018-04-20 paper/related.tex
fixes and acknowledgement

In this paper, we have analyzed real-time performance of a real-world
CNN, which was used in NVIDIA's DAVE-2 self-driving
car~\cite{Bojarski2016}, on a low-cost raspberry pi 3 quad-core 
platform and other embedded multicore platforms. It should be noted, 
however, that DAVE-2's CNN is relatively small compared
to recent state-of-the-art CNNs, which are increasingly larger and
deeper. For example, the CNN based object detector models evaluated in
Google's recent study~\cite{huang2017speed} have between 3M to 54M
parameters, and are much larger than DAVE-2's CNN.
Using such large CNN models will be challenging on
resource constrainted embedded computing platforms, especially for
real-time applications such as self-driving cars.

>>>>

In this paper, we have analyzed real-time performance of a real-world
CNN, which was used in NVIDIA's DAVE-2 self-driving
car~\cite{Bojarski2016}, on a low-cost raspberry pi 3 quad-core 
platform and other embedded multicore platforms. It should be noted, 
however, that DAVE-2's CNN is relatively small compared
to recent state-of-the-art CNNs, which are increasingly larger and
deeper. For example, the CNN based object detector models evaluated in
Google's recent study~\cite{huang2017speed} have between 3M to 54M
parameters, which are much larger than DAVE-2's CNN.
Using such large CNN models will be challenging on
resource constrainted embedded computing platforms, especially for
real-time applications such as self-driving cars.
